\section{Introduction}

Quantum computing holds the potential to efficiently solve important problems that are too complex to tackle on classical computers.  For certain classes of problems, quantum algorithms offer an exponential speedup over their best-known classical counterparts.  Famous examples that illustrate the theoretical advantages of quantum computing include Grover's algorithm for searching through large unstructured databases \cite{grover1996fast} and Shor's algorithm for factoring large semi-prime numbers \cite{Shor_1997}.

However, today's quantum computing hardware is said to exist in the Noisy Intermediate Scale Quantum (NISQ) era, and it is still somewhat of an open question if any advantage over classical computing can be demonstrated on NISQ computers.  This is because current quantum computers are extremely sensitive to environmental perturbations, and most quantum algorithms have much higher resource requirements in terms of qubit count and fidelity than what is available in the near term.  Such perturbations are referred to as noise, and they cause decoherence of the fragile state of the quantum computer.  The NISQ era represents a crucial phase in quantum computing, where although quantum processors are still susceptible to noise and decoherence, they are also too large to be simulated classically. In this context, quantum algorithms tailored for NISQ devices have garnered significant attention. These algorithms aim to strike a delicate balance between harnessing quantum advantages and mitigating the impact of noise, making them practical for today's quantum hardware.  There is no guarantee that quantum computers will ever be stable enough to run algorithms such as those invented by Grover or Shor, so the development of powerful NISQ algorithms is crucial for demonstrating practical utility with quantum computers.

Among the diverse array of NISQ era algorithms, variational quantum algorithms have emerged as promising candidates. These algorithms combine the strengths of both quantum and classical computing, creating a hybrid approach that leverages quantum resources for specific tasks while employing classical methods for everything else.  By restricting usage of quantum processing to the classical computational bottlenecks within a hybrid algorithm, variational quantum algorithms can navigate the challenges posed by noise and decoherence, offering a path towards the realization of near-term practical quantum computations.  

This paper introduces one such variational quantum algorithm called the Variational Quantum Eigensolver (VQE).  VQE is particularly well-suited for solving eigenvalue problems, making it a versatile tool for applications in quantum chemistry, materials science, and optimization.  We discuss the original proposal of the VQE algorithm by Peruzzo et al. in 2014, as well as two recent VQE advancements that extend the applicability of the VQE framework.
\input{Mathematical Background}
\input{lit review}